MLH1 Data Report

### 2019-11-13 doc/Main_Report_Results.rmd
added DMC1 results to results page

@@

@@

### Difference in proportion of chrm classes

>>>>

## MLH1 means

figure 1
mixed models (ummary of mixed model, justificiation for going into post hoc

@@

### Mixed Model for MLH1 counts

### Sex Specific Evolution

#### Male Patterns

rapid evolution in musculus and molossinus; high and low CO count males

#### Female Patterns

seems like female specific evolution in molossinus and musculus.

more variation in domesticus females

## Within Mouse Variation in MLH1 counts

#DMC1 estimates

to localize the evolution in the pathway

  figure
  ANOVA for L and Z
  t-tests betwee high and low groups

```{r DMC1_block1, echo=FALSE}
#loadt DMC1 csv

DMC1.DF <- read.csv("~./DMC1/newJuviQuant.csv", header = TRUE)
DMC1.DF$fileName <- DMC1.DF$Original.Name
DMC1.DF <- add_mouse(DMC1.DF)
DMC1.DF <- add_strain(DMC1.DF)
DMC1.DF <- add_subsp(DMC1.DF)
DMC1.DF <- DMC1.DF[DMC1.DF$stage != "P",]

#1. ANOVA, is there a difference in DMC1 counts across strains?
mod.L <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "L",])
mod.Z <- lm(total.count ~ strain, data = DMC1.DF[DMC1.DF$stage == "Z",])

l.anova <- anova(mod.L) #0.0002701 ***
z.anova <- anova(mod.Z) #NS

l.anova
z.anova
```

In ANOVA analysis strain is significant for Leptotene cells but not Z cells.

```{r DMC1_t.tests, echo=FALSE}
#2. t-tests for the DMC1 counts 
#need to account for sample sizes when comparing t.tests (KAZ and G don't have equal across )
#best case scnario -- more microscoping to get ~20 cells for each stage by strain combo

#L vs Z (all species)
#jj <- t.test(DMC1.male.data$total.count[DMC1.male.data$stage == "Z"], DMC1.male.data$total.count[DMC1.male.data$stage == "L"] )
#p-value = 1.032e-05
#not a valid interpertation

#high vs low group (L)
zola <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
# p-value =  0.002145

#high vs low group (Z)
jesus <- t.test(DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ],
               DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]    )
# p-value = 0.6628

#dom vs musc (L) (2 subsp vs)
#cheer <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "L")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "L")] )
# p-value = 0.003016

#dom vs musc (Z)
#up <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], #DMC1.male.data$total.count[(DMC1.male.data$subsp == "Musc") & (DMC1.male.data$stage == "Z")] )
#p-value = 0.09588

#KAZ vs Dom (Z) (only 1 L cell)
#dont <- t.test(DMC1.male.data$total.count[(DMC1.male.data$subsp == "Dom") & (DMC1.male.data$stage == "Z")], 
#      DMC1.male.data$total.count[(DMC1.male.data$strain == "KAZ") & (DMC1.male.data$stage == "Z")] )
#p-value = 0.004148
#Kaz Z are higher than Dom
```

The p values from the t.tests of the high vs low groups, indecate that the high recombining are significantly higher for the L cells (pvalue = `r zola$p.value`) while the Z cells, there is not a significant difference across the high and low groups (pvalue = `r jesus$p.value`).

```{r DMC1.subsampling, echo=FALSE}

#recreate the DMC1 data by subsampling ~9 or lowest cell number from 
# the 2 groups

#make the 2 pools
High.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]
High.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "MSM") | (DMC1.DF$strain == "PWD") ) )  ]

Low.L.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "L") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]
Low.Z.pool <- DMC1.DF$total.count[( (DMC1.DF$stage == "Z") & ( ( DMC1.DF$strain == "WSB") | (DMC1.DF$strain == "G") | (DMC1.DF$strain == "KAZ") ) )]

#permute drawing / subsampling and performing the t-test
#put the p.value results into a DF!

#permut.SC.cv.diff  <- as.data.frame(
#sample(THE.DF, SAMPLE.SIZE, REPLACE?)

carry.on <-   replicate(10, 
 sample(High.L.pool, 25, replace = FALSE) )#this makes 10x9 DF from the samples

bunchOpvals.L <- replicate(10, 
    # t.test
    t.test( sample(High.L.pool, 25, replace = FALSE),
            sample(Low.L.pool, 25, replace = FALSE), var.equal = FALSE )$p.value
          )

bunchOpvals.Z <- replicate(10, 
    # t.test
    t.test( sample(High.Z.pool, 25, replace = FALSE),
            sample(Low.Z.pool, 25, replace = FALSE), var.equal = FALSE)$p.value
          )

colnames(permut.SC.cv.diff) <- "permuted.diff"

#ASSUMPTIONS; normality, equal variance

#The t.test( ) function produces a variety of t-tests. Unlike most statistical packages, the default assumes #unequal variance and applies the Welsh df modification.
#n use the var.equal = TRUE
#https://www.statmethods.net/stats/ttest.html
```

```{r DMC1.plot, echo=FALSE}
#think of better color scheme
#make the jitter col thinner or make the stacked dots.
#remember that box-plots add outlier dots.
#change y labels

library(ggpubr)
#https://www.r-bloggers.com/add-p-values-and-significance-levels-to-ggplots/

DMC1.DF$strain<- factor(DMC1.DF$strain, levels =c("G","WSB","KAZ",
              "PWD","MSM"), order=T )
#order the data frame
DMC1.DF <- with(DMC1.DF, DMC1.DF[order(strain),])

#for the test bars
my_comparisons <- list( c("KAZ", "MSM"))#no idea how to adjust

#experimenting with ggboplot
everyday <- ggboxplot(DMC1.DF[DMC1.DF$stage != "P",], x = "strain", y = "total.count",
          color = "stage", palette = "jco",
          add = "jitter", short.panel.labs = FALSE)
  #library gpubr has functions to add in sig/test bars.

DMC1.boxplot <- ggplot(aes(x=strain, y=total.count, color=stage), data=DMC1.DF[DMC1.DF$stage != "P",])+
  geom_boxplot(aes(alpha=.5))+geom_jitter()+ggtitle("DMC1 averages in juvenile mice")+

  #this will add neat bars, but still need to figure out what groups ect it's actually using
  stat_compare_means(comparisons = my_comparisons)

#  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
#        panel.background = element_blank(), axis.line = element_line(colour = "black"))

DMC1.boxplot
```

> At what stage are DMC1 foci nnumbers most correlated with MLH1?

Leptotene is the max number of DMC1 -- so this would have more power to capture the variation in MLH1? Look into what baier et al do.

## Heterochiasmy

- Xy xx  (female vs male adjustment)
-	Chrm proportions (

@@

### Proportion of chrm classes
